---
title: "STA_445_Assignment_6"
author: "Sophia Kubisiak"
date: "03/28/24"
output: pdf_document
---

```{r warning=FALSE , message=FALSE}
library(tidyverse)
library(lubridate)
```

## Problem 1
Convert the following to date or date/time objects.

a. September 13, 2010.
```{r}
a <- "September 13, 2010"
mdy(a)
```

b. Sept 13, 2010.
```{r}
b <- "Sept 13, 2010"
mdy(b)
```

c. Sep 13, 2010.
```{r}
c <- "Sep 13, 2010"
mdy(c)
```

d. S 13, 2010. Comment on the month abbreviation needs.
```{r}
d <- "S 13, 2010"
mdy(d)
```
The abbreviations that work are september and sep. Sept and S do not work. 

e. 07-Dec-1941.
```{r}
e <- "07-Dec-1941"
dmy(e)
```

f. 1-5-1998. Comment on why you might be wrong.
```{r}
mdy("1-5-1998")
```
The date could be 'day month year' or 'month day year', it is not clear.   

g. 21-5-1998. Comment on why you know you are correct.
```{r}
dmy("21-5-1998")
```
The date must be read as 'day month year' , because there are not 21 months. 

h. 2020-May-5 10:30 am
```{r}
ymd_hm("2020-May-5, 10:30 AM")
```

i. 2020-May-5 10:30 am PDT (ex Seattle)
```{r}
ymd_hm("2020-May-5 10:30 AM" , tz="US/Pacific")
```

j. 2020-May-5 10:30 am AST (ex Puerto Rico)
```{r}
ymd_hm("2020-May-5 10:30 am" , tz = "America/Puerto_Rico")
```


### Problem 2

Using just your date of birth (ex Sep 7, 1998) and today's date calculate the following:
```{r}
bday <- "June 21, 2002"
birthday <- mdy(bday)
```

a. Calculate the date of your 64th birthday.
```{r}
birthday + years(64)
```

b. Calculate your current age (in years). 
```{r}
as.duration(birthday %--% now())
```

c. Using your result in part (b), calculate the date of your next birthday.
```{r}
(nextbday <- birthday + years(22))
```

d. The number of _days_ until your next birthday.
```{r}
as.period(nextbday %--% now(), unit="days")
```

e. The number of _months_ and _days_ until your next birthday.
```{r}
as.period(nextbday %--% now(), unit="months")
```


## Problem 3

Suppose you have arranged for a phone call to be at 3 pm on May 8, 2015 at Arizona time. However, the recipient will be in Auckland, NZ. What time will it be there? 
```{r}
phonecall <- (mdy_h("May 8, 2015 3PM" , tz="US/Arizona"))
(Auckland <- with_tz(phonecall , tz = "NZ"))
```

## Problem 4

It turns out there is some interesting periodicity regarding the number of births on particular days of the year.

a. Using the `mosaicData` package, load the data set `Births78` which records the number of children born on each day in the United States in 1978. Because this problem is intended to show how to calculate the information using the `date`, remove all the columns *except* `date` and `births`. 

```{r}
data(Births78, package="mosaicData")

births <- Births78 %>%
  select(c("date", "births"))
```

b. Graph the number of `births` vs the `date` with date on the x-axis. What stands out to you? Why do you think we have this trend?

  There are a lot more births between July and October. Maybe this has to do with trends of pregnancy, and possibly marriage dates could have a small impact as well. Also, maybe the amount of planned pregnancies all wanted to give birth in later summer months and early fall in 1978.    
```{r}
ggplot( data=births, aes(x=date, y=births) ) +  
    geom_point(  ) 
```
c. To test your assumption, we need to figure out the what day of the week each observation is. Use `dplyr::mutate` to add a new column named `dow` that is the day of the week (Monday, Tuesday, etc). This calculation will involve some function in the `lubridate` package and the `date` column.
```{r}
births <- births %>%
  mutate(dow = wday(date, label=TRUE))
```

d. Plot the data with the point color being determined by the day of the week variable.
```{r}
ggplot( data=births, aes(x=date, y=births) ) +  
    geom_point( aes(color=dow) ) 
```










